---
title: "Doing Data Science (EDA - Chapter2)"
output: html_document
author: Bijan Ranjbar-Sahraei
---

Exploratary Data Analysis of sample datasets.

# Initialization 
First let's clean up memory
```{r}
rm(list=ls())  
```

This is the Address to data folder
```{r}
data_folder <- "/Users/bian/sandbox/doing_data_science-master/dds_datasets/dds_ch2_nyt/"
```

## Let's play around a bit 
reading the csv file
```{r}
data1 <- read.csv(paste(data_folder, "nyt1.csv", sep=''))
```

To see how data looks like at the begining and at the end
```{r}
head(data1)
tail(data1)
```

To get the column names
```{r}
colnames(data1)
```

To get a summary
```{r}
summary(data1)
```
get a specific columns
```{r}
head(data1$Age)
```

# Ex1. 
Add a new variable age_group that categorizes users between differen age groups
```{r}
data1$age_group <- cut(data1$Age, c(-Inf, 0, 18, 24, 34, 44, 54, 65, Inf))
head(data1)
```

Just to check subset of data which doesn't have any age
```{r}
head(subset(data1, age_group == "(-Inf,0]"))  # it seems that these guys are not sign in, that's why!
```

## doBy
Let's use the "doBy"" package
To do something on data stratified 'by' some other variables. here we use the summaryBy function
```{r}
# install.packages("doBy")  
library("doBy")  
siterange <- function(x) {c(length(x), min(x), mean(x), max(x)) }
siterange(data1$Age)
summaryBy(Age~age_group, data = data1, FUN=siterange)  # summariszing Age based on age-group 
```


Let's see a summary of everything for each age category
```{r}
summaryBy(Gender+Signed_In+Impressions+Clicks~age_group, data=data1)
```
*It seems that the just those who haven't signed in has a zero age.*


# Ex2.a 
Plot the distribution of number impressions and click through-rate for the age categories 

## Plotting Impressions based on age category
let's plot a bit
```{r}
# install.packages("ggplot2")
library(ggplot2)
ggplot(data1, aes(x=Impressions, fill=age_group)) + geom_histogram(binwidth=1)  # woohoo!!
ggplot(data1, aes(x=age_group, y=Impressions, fill=age_group)) + geom_boxplot()

```

Let's check data with no impression
```{r}
head(summaryBy(Clicks~Impressions, data=data1, FUN=siterange))
```
there are 3066 users with no click where the impression is also zero. So, let's remove them.


## Click Through Rate (CTR)
CTR is the number of clicks that your ad receives divided by the number of times your ad is shown expressed as a percentage (clicks / impressions = CTR).

```{r}
data1$hasimps <- cut(data1$Impressions, c(-Inf, 0, Inf))
summaryBy(Clicks~hasimps, data=data1, FUN=siterange)
ggplot(subset(data1,Impressions>0), aes(x=Clicks/Impressions, color=age_group)) + geom_density()
ggplot(subset(data1,Clicks>0 & Age>0), aes(x=Clicks/Impressions, color=age_group)) + geom_density()
```

Let's categorize users based on their click behavior
```{r}
data1$scode[data1$Impressions == 0] <- "NoImps"
data1$scode[data1$Impressions > 0] <- "Imps"
data1$scode[data1$Clicks > 0] <- "Clicks"
```

convert a column to a factor
```{r}
data1$scode <- factor(data1$scode)
```

look at levels
```{r}
clen <- function(x){c(length(x))}
etable <- summaryBy(Impressions~scode+Gender+age_group, data=data1, FUN=clen)
etable
```

# Let's do something serious
Let's make a cleaned version of data by leaving out the users who are not signed in and also the users who don't have any impressions.

```{r}
data_cleaned <- subset(data1, Age>0, Impressions>0)
```

The we add the CTR to the columns
```{r}
data_cleaned <- transform(data_cleaned, CTR = Clicks/Impressions)
```

Let's make a nice plot!
```{r}
# qplot(y=CTR, x=Age, data=data_cleaned, geom=c("boxplot", "jitter"))
```
